---
title: "analysis_bill_tracker"
author: "Laurel Denyer"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
This file conducts descriptive analyses of the bills tracked from an output created in clean_bill_tracker.R. Run that file first to have the proper input for this .Rmd
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = '/Users/laureldenyer/Documents/GitHub/thesis/ca-alternative-cesci') #Change your working directory as needed.
```

Libraries
```{r message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(sf)
library(ggplot2)
library(knitr)
library(DT)
library(RColorBrewer)
library(ggpubr)
library(ggalluvial)
library(gridExtra)
library(gt)
library(ggupset)
library(UpSetR)
library(ComplexUpset)
library(showtext)
library(webshot2)
library(janitor)

```


Visual aesthetics
```{r}

custom_palette <- colorRampPalette(brewer.pal(9, "Set1"))(18)
leg_fate_colors <- c("Enacted" = "#00BF7D", 
                       "Dormant" = "#A3A500", 
                       "Vetoed" = "#E76BF3", 
                       "Failed" = "#00B0F6",
                       "Died" = "#F8766D" )
```

# Data preparation of dummy variables. 
```{r}
bills_dv <- read.csv("data/interim/policy/combined_dv.csv")

bills_dv = bills_dv[-c(1)]   

#grouping the duplicate bills created for DVs. Only used for creating the dvs_only table and cleaning variable names
grouped <- bills_dv %>% 
  # Rename Status columns
  rename_with(~str_replace(.,"^Status","Status_"), starts_with("Status")) %>%  
  rename_with(~str_replace_all(., "(^Topiccat\\.|^Status\\.|\\.)", "_"), matches("^Topic_category_|^Status_")) %>% 
  rename_with(~ str_replace_all(., "\\.", "_") %>% str_replace_all("_{2,}", "_")) %>%
  
  # Remove trailing underscores
  rename_with(~str_replace(., "_$", "")) %>% 
  
  group_by(bill_id) %>% 
  summarise(across(everything(), max, na.rm = TRUE))

# Check column names
colnames(grouped)

#only dvs
dvs_only <- grouped[,c(1,45:285)]
head(dvs_only)

#amended out only
amend_free_dvs <- dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==0, na.rm=TRUE) 

```

#Pulling out the ones with amended out text. Note that the "quick" look sums up to 10, which means 16 are coded as only Amended out.
```{r}
amended_out_bills <- dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1, na.rm=TRUE) %>% 
  select(where(~any(.!=0)))
print(amended_out_bills)

amended_out_bills_quick<- dvs_only %>% 
  dplyr::filter(Status_Amended_out_of_the_bill==1, na.rm=TRUE) %>% 
  dplyr::summarize(
    Enacted = sum(Status_Enacted, na.rm = TRUE),
    Vetoed = sum(Status_Vetoed, na.rm = TRUE),
    Failed = sum(Status_Failed, na.rm = TRUE),
    Dormant = sum(Status_Filed, Status_Held_in_Committee, Status_Inactive_File, Status_Referred, Status_Returned_to_Secretary,Status_Suspended, Status_Introduced, na.rm=TRUE),
    .groups = 'drop'
  )
print(amended_out_bills_quick)

```


### Descriptive statistics

#Numbers of bills over time by legislative session and by fate of the bill
```{r}
legfate_over_time <- dvs_only %>%
  filter(leg_session %in% c("2019-2020", "2021-2022", "2023-2024")) %>%
  dplyr::group_by(leg_session) %>%
  dplyr::summarize(
    Amended_Out = sum(Status_Amended_out_of_the_bill, na.rm = TRUE),
    Enacted = sum(Status_Enacted, na.rm = TRUE),
    Vetoed = sum(Status_Vetoed, na.rm = TRUE),
    Failed = sum(Status_Failed, na.rm = TRUE),
    Died = sum(Status_Died, na.rm = TRUE),
    Dormant = sum(Status_Filed, Status_Held_in_Committee, Status_Inactive_File, Status_Referred, Status_Returned, Status_Returned_to_Secretary,Status_Suspended, Status_Introduced, na.rm=TRUE),
    .groups = 'drop'
  )

print(legfate_over_time)

legfate_over_time_amend_out <- amended_out_bills %>% 
  filter(leg_session %in% c("2019-2020", "2021-2022", "2023-2024")) %>%
  dplyr::group_by(leg_session) %>%
  dplyr::summarize(
    Enacted = sum(Status_Enacted, na.rm = TRUE),
    Vetoed = sum(Status_Vetoed, na.rm = TRUE),
    Dormant = sum(Status_Filed, Status_Held_in_Committee, Status_Inactive_File, na.rm=TRUE),
    .groups = 'drop'
  )
print(legfate_over_time_amend_out)

```
```{r}
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Enacted==1) %>% nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Failed==1) %>% nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Filed==1) %>%  nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Held_in_Committee==1) %>% nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Inactive_File==1) %>% nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Referred==1) %>%  nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Returned_to_Secretary==1) %>%  nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Returned==1) %>% nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Suspended==1) %>%  nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Vetoed==1) %>%  nrow()
dvs_only %>% 
  filter(Status_Amended_out_of_the_bill==1 & Status_Died==1) %>%  nrow()

dvs_only %>%
  filter(Status_Amended_out_of_the_bill == 1) %>%
  group_by(leg_session) %>%
  summarise(
    Died_Count = sum(Status_Died ==1),
    Enacted_Count = sum(Status_Enacted == 1),
    Failed_Count = sum(Status_Failed == 1),
    Filed_Count = sum(Status_Filed == 1),
    Held_in_Committee_Count = sum(Status_Held_in_Committee == 1),
    Inactive_File_Count = sum(Status_Inactive_File == 1),
    Referred_Count = sum(Status_Referred == 1),
    Returned_Count = sum(Status_Returned==1),
    Returned_to_Secretary_Count = sum(Status_Returned_to_Secretary == 1),
    Suspended_Count = sum(Status_Suspended == 1),
    Vetoed_Count = sum(Status_Vetoed == 1)
  )
```



Alluvial visual
```{r pressure, echo=FALSE}
legfate_long <- legfate_over_time %>%
  pivot_longer(cols = c(#Amended_Out, 
    Enacted, Vetoed, Failed, #Introduced, 
    Dormant, Died),
               names_to = "Status",
               values_to = "Count")
print(legfate_long)


alluvial <- ggplot(legfate_long, aes(axis1 = leg_session, axis2 = Status, y = Count)) +
  geom_alluvium(aes(fill = Status),alpha=0.4) +  # Adding the flow of statuses
  geom_stratum(aes(fill = Status)) +  # The blocks representing each status
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +  # Add labels inside the blocks
  scale_x_discrete(limits = c("2019-2020", "2021-2022", "2023-2024"),
                   expand = c(0.15, 0.15)) +
  scale_y_continuous(expand = c(0.05, 0.05)) +
  scale_fill_manual(values = custom_palette) +
  theme_void() +
  labs(title = "Legislative fates of bills",
       x = "Legislative Session",
       y = "Count of Bills",
       fill = "Legislative Fate")


#print(alluvial)
n_legfate <- legfate_long %>% summarise(total = sum(Count, na.rm = TRUE)) %>%
  pull(total)

legfate_bar <- legfate_long %>%
  filter(Count > 0) %>% 
  ggplot(aes(x = leg_session, y = Count, fill = Status)) + # Removed alpha from here
  geom_bar(stat = "identity", position = "stack", color = "black", alpha = 0.7) + # Added alpha globally here
  geom_text(aes(label = Count), position = position_stack(vjust = 0.5), size = 4, color = "white") + 
  scale_fill_manual(values = leg_fate_colors) + # Use scale_fill_manual for custom colors
  theme_minimal() + 
  labs(
    title = paste("Legislative fates of bills, n =", n_legfate),
    x = "Legislative Session",
    y = "Count of Bills",
    fill = "Legislative Fate"
  ) +
  theme_minimal()

  
print(legfate_bar)


ggsave("output/figures/legfate.png", plot=legfate_bar, width=8, height=6, units="in", dpi=300)

#ggsave("output/figures/alluvial.png", plot = alluvial, width = 8, height = 6, units = "in", dpi = 300)
```



#Summary of the ways that CalEnviroScreen and/or Health and Safety Code 39711 is used overall
```{r}
summary_uses <- dvs_only %>%
  select(bill_id, contains("use")) %>% 
  mutate(across(starts_with("I"), ~ifelse(. == 1, 1, 0))) %>%  
  summarise(across(starts_with("I"), sum, na.rm = TRUE)) %>%  
  pivot_longer(cols = starts_with("I"), names_to = "Use", values_to = "use_count1") %>% 
  filter(use_count1 > 0) %>%  
  mutate(Use = gsub("^I[0-9]+_use", "", Use), 
         Use = gsub("_", " ", Use)) %>% 
  group_by(Use) %>% 
  summarise(use_count1 = sum(use_count1), .groups = 'drop') %>% 
  arrange(desc(use_count1))  
print(summary_uses)

summary_uses_amend_free <- amend_free_dvs %>%
  select(bill_id, contains("use")) %>% 
  mutate(across(starts_with("I"), ~ifelse(. == 1, 1, 0))) %>%  
  summarise(across(starts_with("I"), sum, na.rm = TRUE)) %>%  
  pivot_longer(cols = starts_with("I"), names_to = "Use", values_to = "use_count2") %>% 
  filter(use_count2 > 0) %>%  
  mutate(Use = gsub("^I[0-9]+_use", "", Use), 
         Use = gsub("_", " ", Use)) %>% 
  group_by(Use) %>% 
  summarise(use_count2 = sum(use_count2), .groups = 'drop') %>% 
  arrange(desc(use_count2))  
print(summary_uses_amend_free)

merged_summary_uses <- merge(summary_uses, summary_uses_amend_free, by="Use")
merged_summary_uses$Amended_out <- merged_summary_uses$use_count1 - merged_summary_uses$use_count2
merged_summary_uses <- merged_summary_uses %>% arrange(desc(Amended_out))
print(merged_summary_uses)
```
This shows the amended out ones and how they use it, and it follows a similar pattern in that its the most common uses

#Uses by status
```{r}
uses_by_status <- dvs_only %>%
 select(bill_id, leg_session, contains("use"), starts_with("Status_")) %>% 
  mutate(across(starts_with("I"), ~ ifelse(. == 1, 1, 0))) %>%
  pivot_longer(cols = starts_with("I") & contains("use"), names_to = "Use", values_to = "use_count") %>%
  filter(use_count == 1) %>%
  mutate(Use = gsub("^I[0-9]+_use", "", Use),
         Use = gsub("_", " ", Use)) %>%
  pivot_longer(cols = starts_with("Status_"), names_to = "Status", values_to = "status_value") %>%
  filter(!is.na("Status")) %>%
  filter(status_value == 1) %>%
  mutate(Status = case_when(
     Status %in% c("Status_Enacted") ~ "Enacted",
    Status %in% c("Status_Vetoed") ~ "Vetoed",
    Status %in% c("Status_Failed") ~ "Failed",
    Status %in% c("Status_Died") ~ "Died",
    Status %in% c("Status_Filed", "Status_Held_in_Committee", "Status_Inactive_File", "Status_Referred", "Status_Introduced", "Status_Returned_to_Secretary", "Status_Suspended") ~ "Dormant",
  )) %>%
  group_by(leg_session, Use, Status) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(leg_session, Use, desc(use_count))

print(uses_by_status)

uses_by_status_amend_free <- amend_free_dvs %>%
 select(bill_id, leg_session, contains("use"), starts_with("Status_")) %>% 
  mutate(across(starts_with("I"), ~ ifelse(. == 1, 1, 0))) %>%
  pivot_longer(cols = starts_with("I") & contains("use"), names_to = "Use", values_to = "use_count") %>%
  filter(use_count == 1) %>%
  mutate(Use = gsub("^I[0-9]+_use", "", Use),
         Use = gsub("_", " ", Use)) %>%
  pivot_longer(cols = starts_with("Status_"), names_to = "Status", values_to = "status_value") %>%
  filter(status_value == 1) %>%
  mutate(Status = case_when(
    Status %in% c("Status_Enacted") ~ "Enacted",
    Status %in% c("Status_Vetoed") ~ "Vetoed",
    Status %in% c("Status_Failed") ~ "Failed",
    Status %in% c("Status_Died") ~ "Died",
    Status %in% c("Status_Filed", "Status_Held_in_Committee", "Status_Inactive_File", "Status_Referred", "Status_Introduced", "Status_Returned_to_Secretary", "Status_Suspended") ~ "Dormant",
  )) %>%
  group_by(leg_session, Use, Status) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(leg_session, Use, desc(use_count))

print(uses_by_status_amend_free)

#THIS IS NOT WORKING - just going to handle manually 
uses_by_status_amend_out <- amended_out_bills %>%
  select(bill_id, leg_session, contains("use"), starts_with("Status_")) %>% 
  mutate(across(starts_with("I"), ~ ifelse(. == 1, 1, 0))) %>%
   mutate(across(starts_with("Status_"), ~ replace_na(., 0))) %>%
  pivot_longer(cols = starts_with("I") & contains("use"), names_to = "Use", values_to = "use_count") %>%
  filter(use_count == 1) %>%
  mutate(Use = gsub("^I[0-9]+_use", "", Use),
         Use = gsub("_", " ", Use)) %>%
  pivot_longer(cols = starts_with("Status_"), names_to = "Status", values_to = "status_value") %>%
  filter(status_value == 1) %>%
  mutate(Status = case_when(
    Status %in% c("Status_Enacted", "Status_Vetoed", "Status_Failed", "Status_Introduced", "Status_Died") ~ Status,
    Status %in% c("Status_Filed", "Status_Held_in_Committee", "Status_Inactive_File", "Status_Referred", "Status_Returned_to_Secretary", "Status_Suspended") ~ "Dormant",
  )) %>%
  group_by(leg_session, Use, Status) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(leg_session, Use, desc(use_count))
print(uses_by_status_amend_out)

```


More visualizations of use by status
```{r, echo=FALSE}
uses_by_status_plot <- ggplot(uses_by_status, aes(x = Use, y = use_count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use by Status Count",
    x = "Use Type",
    y = "Status Count",
    fill = "Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("output/figures/uses_by_status_plot.png", plot = uses_by_status_plot, width = 8, height = 6, units = "in", dpi = 300)

table_uses_by_status <- uses_by_status %>%
  group_by(Use, Status) %>%
  summarise(use_count = sum(use_count), .groups = 'drop') %>%
  arrange(Use, Status)

uses_point_plot <- ggplot(table_uses_by_status, aes(x = Use, y = Status, fill = use_count, color = use_count)) +
  geom_point(aes(size = use_count), shape = 21) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  scale_color_gradient(low = "lightblue", high = "darkblue") +
  theme_void() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

ggsave("output/figures/uses_point_plot.png", plot = uses_point_plot, width = 8, height = 6, units = "in", dpi = 300)

table_uses_by_status_nested <- table_uses_by_status %>%
  group_by(Use) %>%
  pivot_wider(names_from = Status, values_from = use_count, values_fill = list(use_count = 0)) %>%
  ungroup()

table_uses_by_status_nested_transposed <- table_uses_by_status_nested %>%
  pivot_longer(
    cols = -Use,
    names_to = "Status",
    values_to = "use_count"
  ) %>%
  pivot_wider(
    id_cols = Status,
    names_from = Use,
    values_from = use_count,
    values_fill = list(use_count = 0)
  )

table_uses_by_status_nested_transposed <- table_uses_by_status_nested_transposed %>% 
  mutate(Total = rowSums(.[, sapply(., is.numeric)]))

use_status_totals_row <- data.frame(
  Employ_Type = "Total",
  `Advisory role` = sum(table_uses_by_status_nested_transposed$`Advisory role`, na.rm = TRUE),
  `Community engagement` = sum(table_uses_by_status_nested_transposed$`Community engagement`, na.rm = TRUE),
  `Defining a community` = sum(table_uses_by_status_nested_transposed$`Defining a community`, na.rm = TRUE),
  `Funding prioritization` = sum(table_uses_by_status_nested_transposed$`Funding prioritization`, na.rm = TRUE),
  `Mitigation` = sum(table_uses_by_status_nested_transposed$`Mitigation`, na.rm = TRUE),
  `Program eligibility` = sum(table_uses_by_status_nested_transposed$`Program eligibility`, na.rm = TRUE),
  `Program goals` = sum(table_uses_by_status_nested_transposed$`Program goals`, na.rm = TRUE),
  `Program prioritization` = sum(table_uses_by_status_nested_transposed$`Program prioritization`, na.rm = TRUE),
  `Requires CES to be used in analysis` = sum(table_uses_by_status_nested_transposed$`Requires CES to be used in analysis`, na.rm = TRUE),
  `Used as data in support of writing the bill` = sum(table_uses_by_status_nested_transposed$`Used as data in support of writing the bill`, na.rm = TRUE),  
  Total = sum(table_uses_by_status_nested_transposed$Total, na.rm = TRUE)
)

table_use_status_with_totals <- table_uses_by_status_nested_transposed %>%
  adorn_totals(where = "row")

gt_table_status <- table_use_status_with_totals %>%
  filter(Status != "NA") %>%
  gt() %>%
  tab_options(table.width = pct(100)) %>%
  tab_header(
    title = "Use Count by Legislative Status"
  )

print(gt_table_status)

gtsave(gt_table_status, "output/figures/use_status_comparison_table.png")

```

Same visuals as above but without the amended out bills
```{r, echo=FALSE}
ggplot(uses_by_status_amend_free, aes(x = Use, y = use_count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use by Status Count - amended free",
    x = "Use Type",
    y = "Status Count",
    fill = "Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

table_uses_by_status_amend_free <- uses_by_status_amend_free %>%
  group_by(Use, Status) %>%
  summarise(use_count = sum(use_count), .groups = 'drop') %>%
  arrange(Use, Status)

ggplot(table_uses_by_status_amend_free, aes(x = Use, y = Status, fill = use_count, color = use_count)) +
  geom_point(aes(size = use_count), shape = 21) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  scale_color_gradient(low = "lightblue", high = "darkblue") +
  theme_void() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) 

table_uses_by_status_amend_free_nested <- table_uses_by_status_amend_free %>%
  group_by(Use) %>%
  pivot_wider(names_from = Status, values_from = use_count, values_fill = list(use_count = 0)) %>%
  ungroup()

table_uses_by_status_amend_free_nested %>%
  gt() %>%
  tab_header(
    title = "Use Count by Status - amended free"
  )
```

```{r}
upset_uses_by_status <- dvs_only %>%
 select(bill_id, leg_session, contains("use"), starts_with("Status_")) %>% 
  mutate(across(starts_with("I"), ~ ifelse(. == 1, 1, 0))) %>%
  pivot_longer(cols = starts_with("I") & contains("use"), names_to = "Use", values_to = "use_count") %>%
  filter(use_count == 1) %>%
  mutate(Use = gsub("^I[0-9]+_use", "", Use),
         Use = gsub("_", " ", Use)) %>%
  pivot_longer(cols = starts_with("Status_"), names_to = "Status", values_to = "status_value") %>%
  filter(!is.na("Status")) %>%
  filter(status_value == 1) %>%
  mutate(Status = case_when(
    Status %in% c("Status_Enacted") ~ "Enacted",
    Status %in% c("Status_Vetoed") ~ "Vetoed",
    Status %in% c("Status_Failed") ~ "Failed",
    Status %in% c("Status_Died") ~ "Died",
    Status %in% c("Status_Filed", "Status_Held_in_Committee", "Status_Inactive_File", "Status_Referred", "Status_Introduced", "Status_Returned_to_Secretary", "Status_Suspended") ~ "Dormant",
  )) %>%
  group_by(leg_session, Use, Status) 

print(upset_uses_by_status)

upset_uses_by_status <- upset_uses_by_status %>% select(-c(Topic_category_Land_use))

# Pivot wider
wide_data <- upset_uses_by_status %>%
  mutate(use_count = as.numeric(use_count)) %>%  # Ensure numeric type
  pivot_wider(
names_from = Use,
 values_from = use_count,
 values_fill = list(use_count = 0),
 values_fn = list(use_count = sum)  # Handle duplicates
)
wide_data_clean <- na.omit(wide_data)


uses=colnames(wide_data)[5:15]

upset_stack <- upset( 
    wide_data_clean,
    uses, name='Use', width_ratio = 0.3, set_sizes=FALSE, min_size = 2, 
    base_annotations=list(
        'Intersection size'=intersection_size(
            counts=FALSE,
            show_numbers = FALSE,
            mapping=aes(fill=Status))
    )
)+ 
  ggtitle("Intensity of uses within bills, n = 216 bills, 566 instances") +
  scale_fill_manual(values=leg_fate_colors) +
  theme(
    plot.title = element_text(
      hjust = 0.5,   # center the title
      vjust = 2      # move the title higher above the plot
    )
  )
print(upset_stack)

upset_simple <- upset(wide_data_clean, uses, name='Use', width_ratio = 0.3, set_sizes=FALSE, min_size = 2, 
      base_annotations = list(
        'Intersection size' = intersection_size(
            counts = FALSE,  
            show_numbers = FALSE
        )))

print(upset_simple)

ggsave("output/figures/upset_stack.png", plot=upset_stack, width=8, height=6, units="in", dpi=300)

ggsave("output/figures/upset.png", plot=upset_simple, width=8, height=6, units="in", dpi=300)


```


```{r, echo=FALSE}
ggplot(uses_by_status, aes(x = Status, y = use_count, fill = Use)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use Count by Status",
    x = "Status",
    y = "Use Count",
    fill = "Use Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(uses_by_status_amend_free, aes(x = Status, y = use_count, fill = Use)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use Count by Status - amended free",
    x = "Status",
    y = "Use Count",
    fill = "Use Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
Parse out the bills with multi domain assignments
```{r}
multi_domain <- bills_dv %>% 
  group_by(bill_id) %>%
  filter(n_distinct(Topic.category) > 1)

unique(multi_domain$bill_id)
n_distinct(multi_domain$bill_id)

```


#Topics of bills
```{r}
topics <- dvs_only %>%
  select(bill_id, starts_with("Topic_category_")) %>%
  gather(key="Topic_category",value="value",-bill_id) %>% 
  filter(value==1) %>% 
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category_Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_categorysea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste",Topic_category,ignore.case=TRUE)~"Waste",
    TRUE ~ gsub("_"," ", gsub("^Topic_category_","",Topic_category))
  )) %>%
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>% 
  arrange(desc(bill_count)) 
print(topics)

topics_amend_free <- amend_free_dvs %>%
  select(bill_id, starts_with("Topic_category_")) %>%
  gather(key="Topic_category",value="value",-bill_id) %>% 
  filter(value==1) %>% 
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category__Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste",Topic_category,ignore.case=TRUE)~"Waste",
    TRUE ~ gsub("_"," ", gsub("^Topic_category_","",Topic_category))
  )) %>%
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>% 
  arrange(desc(bill_count)) 
print(topics_amend_free)

topics_amend_out <- amended_out_bills %>%
  select(bill_id, starts_with("Topic_category_")) %>%
  gather(key="Topic_category",value="value",-bill_id) %>% 
  filter(value==1) %>% 
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category_Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste",Topic_category,ignore.case=TRUE)~"Waste",
    TRUE ~ gsub("_"," ", gsub("^Topic_category_","",Topic_category))
  )) %>%
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>% 
  arrange(desc(bill_count)) 
print(topics_amend_out)
```

#Removing budget bills
```{r}

dvs_titled <- dvs_only %>%
  left_join(grouped %>% select(bill_id, Topic_Title), by = "bill_id")

dvs_titled_amend_free <- amend_free_dvs %>%
  left_join(grouped %>% select(bill_id, Topic_Title), by = "bill_id")

dvs_titled_amend_out <- amended_out_bills %>%
  left_join(grouped %>% select(bill_id, Topic_Title), by = "bill_id")

topics_without_budget <- dvs_titled %>%
  select(bill_id, starts_with("Topic_category_"),Topic_Title) %>%
  gather(key = "Topic_category", value = "value", -bill_id,-Topic_Title) %>%
  filter(value == 1) %>%
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category_Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste", Topic_category, ignore.case = TRUE) ~ "Waste",
    TRUE ~ gsub("_", " ", gsub("^Topic_category_", "", Topic_category))
  )) %>%
  filter(!grepl("Budget Act", Topic_Title, ignore.case = TRUE)) %>%  
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>%
  arrange(desc(bill_count))

print(topics_without_budget)

topics_without_budget_amend_free <- dvs_titled_amend_free %>%
  select(bill_id, starts_with("Topic_category_"),Topic_Title) %>%
  gather(key = "Topic_category", value = "value", -bill_id,-Topic_Title) %>%
  filter(value == 1) %>%
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category_Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste", Topic_category, ignore.case = TRUE) ~ "Waste",
    TRUE ~ gsub("_", " ", gsub("^Topic_category_", "", Topic_category))
  )) %>%
  filter(!grepl("Budget Act", Topic_Title, ignore.case = TRUE)) %>%  
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>%
  arrange(desc(bill_count))

print(topics_without_budget_amend_free)

topics_without_budget_amend_out <- dvs_titled_amend_out %>%
  select(bill_id, starts_with("Topic_category_"),Topic_Title) %>%
  gather(key = "Topic_category", value = "value", -bill_id,-Topic_Title) %>%
  filter(value == 1) %>%
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",
    grepl("Topic_category_Water", Topic_category, ignore.case = TRUE) ~ "Non-drinking Water",
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards", Topic_category, ignore.case = TRUE) ~ "Climate emergency/natural hazards",
    grepl("Topic_category_Waste", Topic_category, ignore.case = TRUE) ~ "Waste",
    TRUE ~ gsub("_", " ", gsub("^Topic_category_", "", Topic_category))
  )) %>%
  filter(!grepl("Budget Act", Topic_Title, ignore.case = TRUE)) %>%  
  group_by(Topic) %>%
  summarise(bill_count = n(), .groups = 'drop') %>%
  arrange(desc(bill_count))

print(topics_without_budget_amend_out)
```

#Topics by session without budget bills
```{r, echo=FALSE}
topics_by_session <- dvs_titled %>%
  select(bill_id, starts_with("Topic_category_"), Topic_Title, leg_session) %>% 
  filter(leg_session%in% c("2019-2020","2021-2022","2023-2024")) %>% 
  gather(key = "Topic_category", value = "value", -bill_id, -Topic_Title, -leg_session) %>% 
  filter(value == 1) %>%
  mutate(Topic = case_when(
    grepl("^Topic_category_Transportation|^Topic_category_Ports", Topic_category) ~ "Transportation",
    grepl("^Topic_category_Governance", Topic_category) ~ "Governance",#combines governance types
    grepl("Topic_category_Energy|Topic_category_Electricity", Topic_category, ignore.case = TRUE) ~ "Energy",
    grepl("Topic_category_sea_level_rise|Topic_category_Climate_emergency_natural_hazards|Topic_category_Carbon_emissions", Topic_category, ignore.case = TRUE) ~ "Climate",
    grepl("Topic_category_Housing|Topic_category_Greening|Topic_category_Land_use|Topic_category_Planning", Topic_category, ignore.case=TRUE) ~ "Planning",
    grepl("Topic_category_Air_pollution_quality_nonvehicular|Topic_category_Groundwater|Topic_category_Marine|Topic_category_Water", Topic_category, ignore.case=TRUE) ~ "Environmental conditions",
    grepl("Topic_category_Oil_gas", Topic_category, ignore.case = TRUE) & bill_id %in% c("95", "146") ~ "Planning",
    grepl("Topic_category_Oil_gas", Topic_category, ignore.case = TRUE) & bill_id %in% c("162") ~ "Environmental conditions",
    grepl("Topic_category_Oil_gas", Topic_category, ignore.case = TRUE) & bill_id %in% c("168") ~ "Energy",
    grepl("Topic_category_Community_development_grants|Topic_category_Education_and_workforce", Topic_category, ignore.case=TRUE) ~ "Community development",
    grepl("Topic_category_Waste", Topic_category, ignore.case = TRUE) ~ "Waste", #combines waste types
    TRUE ~ gsub("_", " ", gsub("^Topic_category_", "", Topic_category))  
  )) %>%
  filter(!grepl("Budget Act", Topic_Title, ignore.case = TRUE)) %>%  
  group_by(leg_session, Topic) %>%  
  summarise(bill_count = n(), .groups = 'drop') %>%
  pivot_wider(names_from=leg_session, values_from=bill_count, values_fill = 0)
print(topics_by_session)

topics_by_session_long <-topics_by_session %>% 
  pivot_longer(cols=c("2019-2020", "2021-2022", "2023-2024"),
               names_to="leg_session",
               values_to="bill_count")
n_topics <- topics_by_session_long %>% summarise(total = sum(bill_count, na.rm = TRUE)) %>%
  pull(total)

bar_bills_topics_time <- ggplot(topics_by_session_long, aes(x = Topic, y = bill_count, fill = Topic)) +
  geom_bar(stat = "identity", position = "dodge") + 
  facet_wrap(~ leg_session) + 
  labs(
    title = paste("Bill Count by Domain Across Legislative Sessions, n =", n_topics),
    x = NULL,
    y = "Number of bills in each domain",
    fill = "Domain"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(), 
    axis.title.x = element_blank() 
  )

print(bar_bills_topics_time)
ggsave("output/figures/bar_bills_topics_time.png", plot = bar_bills_topics_time, width = 8, height = 6, units = "in", dpi = 300)
```


## Adding and amending of codes
# Action type (added or amended to code) by bill
```{r, echo=FALSE}
action_summary <- dvs_only %>%
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )
summary(action_summary)
action_summary %>% 
  #filter(Added>0) %>% 
  ggplot(aes(x = Added)) +
  geom_histogram() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1)
action_summary %>% 
  #filter(Added>0) %>% 
  ggplot(aes(x = Amended)) +
  geom_histogram() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1)


action_summary_amend_free <- amend_free_dvs %>%
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )
action_summary_amend_out <- amended_out_bills %>%
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )

#group by bill id, is it just adding or amending happening or a mix
action_summary <- dvs_only %>% 
  group_by(bill_id) %>% 
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )
print(action_summary)

action_summary_amend_free <- amend_free_dvs %>% 
  group_by(bill_id) %>% 
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )
print(action_summary_amend_free)

action_summary_amend_out <- amended_out_bills %>% 
  group_by(bill_id) %>% 
  summarize(
    Added = sum(across(starts_with("added")), na.rm = TRUE),
    Amended = sum(across(starts_with("amend")), na.rm = TRUE),
    .groups = 'drop'
  )
print(action_summary_amend_out)


action_summary_long <- action_summary %>%
  pivot_longer(cols = c(Added, Amended), 
               names_to = "Action", 
               values_to = "Count")

add_amend_plot <- ggplot(action_summary_long, aes(x = Action, y = Count, color = Action)) +
  geom_point(size = 2, position = position_jitter(width = 0.1)) + 
  labs(title = "Distribution of Added and Amended text in bills",
       x = "Action Type",
       y = "Count per bill") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) 

add_amend_plot <- ggplot(action_summary_long, aes(x = Action, y = Count, color = Action)) +
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE) + 
  labs(title = "Distribution of Added and Amended text in bills",
       x = "Action Type",
       y = "Count per bill") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) 

add_amend_plot <- ggplot(action_summary_long, aes(x = Action, y = Count, color=Action)) +
  geom_boxplot(outlier.shape = NA, fill = "turquoise", color = "black") +  #remove outliers
  geom_jitter(width = 0.2, alpha = 0.2, color = "black") +
  #stat_summary(fun = mean, geom = "point", color = "darkturquoise", size = 3) +
  labs(title ="Distribution of Added and Amended text in bills" ,
       x = "Action Type",
       y = "Count per bill") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(add_amend_plot)
ggsave("output/figures/add_amend.png", plot = add_amend_plot, width = 8, height = 6, units = "in", dpi = 300)

action_summary_amend_free_long <- action_summary_amend_free %>%
  pivot_longer(cols = c(Added, Amended), 
               names_to = "Action", 
               values_to = "Count")

ggplot(action_summary_amend_free_long, aes(x = Action, y = Count, color = Action)) +
  geom_point(size = 3, position = position_jitter(width = 0.1)) + 
  labs(title = "Distribution of Added and Amended Actions - Amended out text gone",
       x = "Action Type",
       y = "Count of Actions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) 

action_summary_amend_out_long <- action_summary_amend_out %>%
  pivot_longer(cols = c(Added, Amended), 
               names_to = "Action", 
               values_to = "Count")

ggplot(action_summary_amend_out_long, aes(x = Action, y = Count, color = Action)) +
  geom_point(size = 3, position = position_jitter(width = 0.1)) + 
  labs(title = "Distribution of Added and Amended Actions - Amended out text only",
       x = "Action Type",
       y = "Count of Actions") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) 
```
#Sections of code being added or amended to
```{r}
location_code_types <- dvs_only %>% 
    summarize(
    "Public utilities" = sum(across(contains("public_utilities")), na.rm = TRUE),
    "Health & safety" = sum(across(contains("health_safety")), na.rm = TRUE),
    "Streets & highways" = sum(across(contains("streets_highways")), na.rm = TRUE),
    "Public resources" = sum(across(contains("public_resources")), na.rm = TRUE),
    "Government" = sum(across(contains("government")), na.rm = TRUE),
    "Food & ag" = sum(across(contains("food_agricultural")), na.rm = TRUE),
    "Education" = sum(across(contains("education")), na.rm = TRUE),
    "Civil" = sum(across(contains("civil")), na.rm = TRUE),
    "Water" = sum(across(contains("water")), na.rm = TRUE),
    "Fish & game" = sum(across(contains("fish_game")), na.rm = TRUE),
    "Statutes" = sum(across(contains("statutes")), na.rm = TRUE),
    "Findings" = sum(across(contains("findings")), na.rm = TRUE),
    .groups = 'drop'
  )
location_code_long <- location_code_types %>%
  pivot_longer(cols = c("Public utilities", "Health & safety", "Streets & highways", "Public resources", "Government", "Food & ag", "Education", "Civil", "Water", "Fish & game", "Statutes", "Findings"), 
               names_to = "Location", 
               values_to = "Count") %>% 
   arrange(desc(Count))

print(location_code_long)

location_code_table <- location_code_long %>%
  gt() %>%
  tab_header(
    title = "Government codes mentioning CES"
  )

# Print the gt table
location_code_table
gtsave(location_code_table, "output/figures/code_table.png")



location_code_types_amend_free <- amend_free_dvs %>% 
    summarize(
    "Public utilities" = sum(across(contains("public_utilities")), na.rm = TRUE),
    "Health & safety" = sum(across(contains("health_safety")), na.rm = TRUE),
    "Streets & highways" = sum(across(contains("streets_highways")), na.rm = TRUE),
    "Public resources" = sum(across(contains("public_resources")), na.rm = TRUE),
    "Government" = sum(across(contains("government")), na.rm = TRUE),
    "Food & agriculture" = sum(across(contains("food_agricultural")), na.rm = TRUE),
    "Education" = sum(across(contains("education")), na.rm = TRUE),
    "Civil" = sum(across(contains("civil")), na.rm = TRUE),
    "Water" = sum(across(contains("water")), na.rm = TRUE),
    "Fish & game" = sum(across(contains("fish_game")), na.rm = TRUE),
    "Statutes" = sum(across(contains("statutes")), na.rm = TRUE),
    "Findings" = sum(across(contains("findings")), na.rm = TRUE),
    .groups = 'drop'
  )
location_code_long_amend_free <- location_code_types_amend_free %>%
  pivot_longer(cols = c(Public_utilities, Health_safety, Streets_highways, Public_resources, Government, Food_ag, Education, Civil, Water, Fish_game, Statutes, Findings), 
               names_to = "Location", 
               values_to = "Count") %>% 
   arrange(desc(Count))

print(location_code_long_amend_free)


location_code_types_amend_out <- amended_out_bills %>% 
    summarize(
    "Public utilities" = sum(across(contains("public_utilities")), na.rm = TRUE),
    "Health safety" = sum(across(contains("health_safety")), na.rm = TRUE),
    "Streets highways" = sum(across(contains("streets_highways")), na.rm = TRUE),
    "Public resources" = sum(across(contains("public_resources")), na.rm = TRUE),
    "Government" = sum(across(contains("government")), na.rm = TRUE),
    "Food ag" = sum(across(contains("food_agricultural")), na.rm = TRUE),
    "Education" = sum(across(contains("education")), na.rm = TRUE),
    "Civil" = sum(across(contains("civil")), na.rm = TRUE),
    "Water" = sum(across(contains("water")), na.rm = TRUE),
    "Fish game" = sum(across(contains("fish_game")), na.rm = TRUE),
    "Statutes" = sum(across(contains("statutes")), na.rm = TRUE),
    "Findings" = sum(across(contains("findings")), na.rm = TRUE),
    .groups = 'drop'
  )
location_code_long_amend_out <- location_code_types_amend_out %>%
  pivot_longer(cols = c(Public_utilities, Health_safety, Streets_highways, Public_resources, Government, Food_ag, Education, Civil, Water, Fish_game, Statutes, Findings), 
               names_to = "Location", 
               values_to = "Count") %>% 
   arrange(desc(Count))

print(location_code_long_amend_out)
```

##Employ of HSC vs CES
```{r}
summary_employ <- dvs_only %>%
  select(bill_id, contains("employ")) %>% 
  mutate(across(starts_with("I"), ~ifelse(. == 1, 1, 0))) %>%  
  summarise(across(starts_with("I"), sum, na.rm = TRUE)) %>%  
  pivot_longer(cols = starts_with("I"), names_to = "Employ", values_to = "employ_count") %>% 
  filter(employ_count > 0) %>%  
  mutate(Employ = gsub("^I[0-9]+_employ", "", Employ), 
         Employ = gsub("_", " ", Employ)) %>% 
  group_by(Employ) %>% 
  summarise(employ_count = sum(employ_count), .groups = 'drop') %>% 
  arrange(desc(employ_count))  

print(summary_employ)

summary_employ_amend_free <- amend_free_dvs %>%
  select(bill_id, contains("employ")) %>% 
  mutate(across(starts_with("I"), ~ifelse(. == 1, 1, 0))) %>%  
  summarise(across(starts_with("I"), sum, na.rm = TRUE)) %>%  
  pivot_longer(cols = starts_with("I"), names_to = "Employ", values_to = "employ_count") %>% 
  filter(employ_count > 0) %>%  
  mutate(Employ = gsub("^I[0-9]+_employ", "", Employ), 
         Employ = gsub("_", " ", Employ)) %>% 
  group_by(Employ) %>% 
  summarise(employ_count = sum(employ_count), .groups = 'drop') %>% 
  arrange(desc(employ_count))  

print(summary_employ_amend_free)

summary_employ_amend_out <- amended_out_bills %>%
  select(bill_id, contains("employ")) %>% 
  mutate(across(starts_with("I"), ~ifelse(. == 1, 1, 0))) %>%  
  summarise(across(starts_with("I"), sum, na.rm = TRUE)) %>%  
  pivot_longer(cols = starts_with("I"), names_to = "Employ", values_to = "employ_count") %>% 
  filter(employ_count > 0) %>%  
  mutate(Employ = gsub("^I[0-9]+_employ", "", Employ), 
         Employ = gsub("_", " ", Employ)) %>% 
  group_by(Employ) %>% 
  summarise(employ_count = sum(employ_count), .groups = 'drop') %>% 
  arrange(desc(employ_count))  

print(summary_employ_amend_out)
```


```{r}
use_employ_comparison <- dvs_only %>%
  select(bill_id, contains("use"), contains("employ")) %>%
  pivot_longer(cols = contains("use"), names_to = "Use_Type", values_to = "Use_Count") %>%
  pivot_longer(cols = contains("employ"), names_to = "Employ_Type", values_to = "Employ_Count") %>%
  filter(Use_Count == 1, Employ_Count == 1) %>%
  mutate(
    Use_Type = gsub("^I[0-9]+_use", "", Use_Type),
    Use_Type = gsub("_", " ", Use_Type),
    Employ_Type = gsub("^I[0-9]+_employ", "", Employ_Type),
    Employ_Type = gsub("_", " ", Employ_Type)
  ) %>%
  group_by(Use_Type, Employ_Type) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(Use_Type, desc(use_count))

print(use_employ_comparison)

use_employ_comparison_flipped <- dvs_only %>%
  select(bill_id, contains("use"), contains("employ")) %>%
  pivot_longer(cols = contains("use"), names_to = "Use_Type", values_to = "Use_Count") %>%
  pivot_longer(cols = contains("employ"), names_to = "Employ_Type", values_to = "Employ_Count") %>%
  filter(Use_Count == 1, Employ_Count == 1) %>%
  mutate(
    Use_Type = gsub("^I[0-9]+_use", "", Use_Type),
    Use_Type = gsub("_", " ", Use_Type),
    Employ_Type = gsub("^I[0-9]+_employ", "", Employ_Type),
    Employ_Type = gsub("_", " ", Employ_Type)
  ) %>%
  group_by(Employ_Type, Use_Type) %>%  # Flip grouping: Employ_Type comes first
  summarise(employ_count = n(), .groups = 'drop') %>%  # Rename count to employ_count
  arrange(Employ_Type, desc(employ_count))  # Sort by Employ_Type and employ_count

# View the flipped data
head(use_employ_comparison_flipped)


use_employ_comparison_amend_free <- amend_free_dvs %>%
  select(bill_id, contains("use"), contains("employ")) %>%
  pivot_longer(cols = contains("use"), names_to = "Use_Type", values_to = "Use_Count") %>%
  pivot_longer(cols = contains("employ"), names_to = "Employ_Type", values_to = "Employ_Count") %>%
  filter(Use_Count == 1, Employ_Count == 1) %>%
  mutate(
    Use_Type = gsub("^I[0-9]+_use", "", Use_Type),
    Use_Type = gsub("_", " ", Use_Type),
    Employ_Type = gsub("^I[0-9]+_employ", "", Employ_Type),
    Employ_Type = gsub("_", " ", Employ_Type)
  ) %>%
  group_by(Use_Type, Employ_Type) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(Use_Type, desc(use_count))

print(use_employ_comparison_amend_free)

use_employ_comparison_amend_out <- amended_out_bills %>%
  select(bill_id, contains("use"), contains("employ")) %>%
  pivot_longer(cols = contains("use"), names_to = "Use_Type", values_to = "Use_Count") %>%
  pivot_longer(cols = contains("employ"), names_to = "Employ_Type", values_to = "Employ_Count") %>%
  filter(Use_Count == 1, Employ_Count == 1) %>%
  mutate(
    Use_Type = gsub("^I[0-9]+_use", "", Use_Type),
    Use_Type = gsub("_", " ", Use_Type),
    Employ_Type = gsub("^I[0-9]+_employ", "", Employ_Type),
    Employ_Type = gsub("_", " ", Employ_Type)
  ) %>%
  group_by(Use_Type, Employ_Type) %>%
  summarise(use_count = n(), .groups = 'drop') %>%
  arrange(Use_Type, desc(use_count))

print(use_employ_comparison_amend_out)

```
#Visualizations of use and employ, just trying out some visualizations
```{r, echo=FALSE}
use_employ_comparison_plot <- ggplot(use_employ_comparison_flipped, aes(x = Use_Type, y = employ_count, fill = Employ_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Employ Categories by Use",
    x = "Use Type",
    y = "Employ Count",
    fill = "Employ Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
print(use_employ_comparison_plot)



ggplot(use_employ_comparison_amend_free, aes(x = Employ_Type, y = use_count, fill = Use_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use Count by Employ Categories - amend free",
    x = "Employ Category",
    y = "Use Count",
    fill = "Use Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

print(use_employ_comparison_plot)


ggplot(use_employ_comparison_amend_out, aes(x = Employ_Type, y = use_count, fill = Use_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Use Count by Employ Categories - amended out",
    x = "Employ Category",
    y = "Use Count",
    fill = "Use Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
ggsave("output/figures/use_employ_comparison_plot.png", plot = use_employ_comparison_plot, width = 8, height = 6, units = "in", dpi = 300)

```
Dot plots
```{r}
use_employ_comparison_dotplot <- ggplot(use_employ_comparison, aes(x = Employ_Type, y = Use_Type, size = use_count, fill = use_count)) +
  geom_point(shape = 21, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_size_continuous(range = c(2, 10)) +
  labs(
    title = "Use  Employ Categories",
    x = "Employ Category",
    y = "Use Category",
    size = "Use Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

ggplot(use_employ_comparison_amend_free, aes(x = Employ_Type, y = Use_Type, size = use_count, fill = use_count)) +
  geom_point(shape = 21, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_size_continuous(range = c(2, 10)) +
  labs(
    title = "Use  Employ Categories",
    x = "Employ Category",
    y = "Use Category",
    size = "Use Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

ggplot(use_employ_comparison_amend_out, aes(x = Employ_Type, y = Use_Type, size = use_count, fill = use_count)) +
  geom_point(shape = 21, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_size_continuous(range = c(2, 10)) +
  labs(
    title = "Use  Employ Categories",
    x = "Employ Category",
    y = "Use Category",
    size = "Use Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
ggsave("output/figures/use_employ_comparison_dotplot.png", plot = use_employ_comparison_dotplot, width = 8, height = 6, units = "in", dpi = 300)


```

Tables
```{r}

table_use_employ_comparison_nested <- use_employ_comparison %>%
  pivot_wider(names_from = Use_Type, values_from = use_count, values_fill = list(use_count = 0))

table_use_employ_comparison_nested <- table_use_employ_comparison_nested %>% 
  mutate(Total = rowSums(.[, sapply(., is.numeric)]))

totals_row <- data.frame(
  Employ_Type = "Total",
  `Advisory role` = sum(table_use_employ_comparison_nested$`Advisory role`, na.rm = TRUE),
  `Community engagement` = sum(table_use_employ_comparison_nested$`Community engagement`, na.rm = TRUE),
  `Defining a community` = sum(table_use_employ_comparison_nested$`Defining a community`, na.rm = TRUE),
  `Funding prioritization` = sum(table_use_employ_comparison_nested$`Funding prioritization`, na.rm = TRUE),
  `Mitigation` = sum(table_use_employ_comparison_nested$`Mitigation`, na.rm = TRUE),
  `Program eligibility` = sum(table_use_employ_comparison_nested$`Program eligibility`, na.rm = TRUE),
  `Program goals` = sum(table_use_employ_comparison_nested$`Program goals`, na.rm = TRUE),
  `Program prioritization` = sum(table_use_employ_comparison_nested$`Program prioritization`, na.rm = TRUE),
  `Requires CES to be used in analysis` = sum(table_use_employ_comparison_nested$`Requires CES to be used in analysis`, na.rm = TRUE),
  `Used as data in support of writing the bill` = sum(table_use_employ_comparison_nested$`Used as data in support of writing the bill`, na.rm = TRUE),  
  Total = sum(table_use_employ_comparison_nested$Total, na.rm = TRUE)
)

table_with_totals <- table_use_employ_comparison_nested %>%
  adorn_totals(where = "row")

gt_table <- table_with_totals %>%
  gt() %>%
  tab_options(table.width = pct(100)) %>%
  tab_header(
    title = "Use Count by Employ Types"
  )

gt_table <- table_with_totals %>%
  select(-"Topic category Land use") %>%
  gt() %>%
  tab_options(
    table.width = pct(100),
 #   column_labels.font.size = 12,
   # row.striping.include_table_body = TRUE,
    column_labels.padding = px(5)
  ) %>%
  tab_header(
    title = "Use Count by Employ Types"
  )


print(gt_table)
gtsave(gt_table, "output/figures/use_employ_comparison_table.png")

table_use_employ_comparison_amend_free_nested <- use_employ_comparison_amend_free %>%
  pivot_wider(names_from = Use_Type, values_from = use_count, values_fill = list(use_count = 0))

table_use_employ_comparison_amend_free_nested %>%
  gt() %>%
  tab_header(
    title = "Use Count by Employ Types - Amended free"
  )



table_use_employ_comparison_amend_out_nested <- use_employ_comparison_amend_out %>%
  pivot_wider(names_from = Use_Type, values_from = use_count, values_fill = list(use_count = 0))

table_use_employ_comparison_amend_out_nested %>%
  gt() %>%
  tab_header(
    title = "Use Count by Employ Types - Amended out only"
  )
```

